---
title: "Using SimpleSVM for Binary Classification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using SimpleSVM for Binary Classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(SimpleSVM)
```

## Introduction
SimpleSVM is an R package that implements a simple support vector machine (SVM) for binary classification using gradient descent. This vignette demonstrates how to use the simple_svm function, provides an example of its output, and compares its performance with the svm() function from the e1071 package.

Example: Training a Simple SVM Model
In this example, we’ll generate some sample data, train the model using simple_svm, and display the results.

``` {r}
# Generate sample data
set.seed(123)
x <- matrix(rnorm(20), ncol = 2)
y <- ifelse(x[,1] + x[,2] > 0, 1, -1)

# Train the simple SVM model
model <- simple_svm(x, y)

# Display the weights and intercept
model$weights
model$intercept
```

## Comparison with svm() from e1071
For comparison, we’ll use the svm() function from the e1071 package, which provides a more advanced SVM implementation.

```{r}
# Install e1071 if not already installed
if (!requireNamespace("e1071", quietly = TRUE)) {
  install.packages("e1071")
}
library(e1071)

e1071_SVM <- svm(x, as.factor(y), type = 'C-classification', kernel = 'linear')

# Extract details
support_vectors <- e1071_SVM$SV
alpha <- e1071_SVM$coefs
intercept <- e1071_SVM$rho
weights <- t(alpha) %*% support_vectors

# Display the weights and intercept
cat("Weights:", weights, "\n")
cat("Intercept:", intercept, "\n")
```


## Performance Comparison
To benchmark the performance of simple_svm versus svm() from e1071, we’ll use the bench package.

```{r}
# Install bench if not already installed
if (!requireNamespace("bench", quietly = TRUE)) {
  install.packages("bench")
}
library(bench)

# Benchmark performance
bench::mark(
  SimpleSVM = { simple_svm(x, y); NULL },
  e1071_SVM = { svm(x, y_factor, type = 'C-classification', kernel = 'linear'); NULL }
)
```

## Conclusion
This vignette demonstrated how to use the simple_svm function for binary classification, including a comparison with the svm() function from e1071. SimpleSVM provides a minimalistic implementation, which is ideal for learning the basics of SVMs and understanding gradient descent.

